---
title: "Exploratory data analysis"
output:
  html_document:
    df_print: paged
---

# Packages

```{r}
library(fs)
library(dplyr)
library(readr)
library(tidyr)
library(innovar) # paquete del lab
library(sf) # simple features: trabajar con data geográfica
library(ggplot2)
theme_set(theme_bw())
library(scales)
library(stplanr) # gráficos de redes en mapas
library(viridis) # colores
```

# Load mobilization data

```{r}
interim_path <- path("data", "interim")
file_name <- "mob-lima-metro-od.csv"
file_path <- path(interim_path, file_name)
od <- read_csv(file_path)
glimpse(od)
```

```{r}
length(unique(od$ubigeo_ori))
length(unique(od$ubigeo_des))
```

# Load population data

```{r}
raw_path <- path("data", "raw")
pop_file_name <- "pop-census-2017.csv"
pop_file_path <- path(raw_path, pop_file_name)
pop <- read_csv(pop_file_path)
glimpse(pop)
```

```{r}
#TODO: Intra-district analysis
```

# Inter-distrital analysis

```{r}
od %>% 
  filter(ubigeo_ori != ubigeo_des) ->
  inter_od
```

## Districts that are not a destiny 

```{r}
length(unique(inter_od$ubigeo_des))
setdiff(od$distr_des, inter_od$distr_des)
```

## Districts that are not an origin 

```{r}
length(unique(inter_od$ubigeo_ori))
setdiff(od$distr_ori, inter_od$distr_ori)
```

## In-flow

```{r}
inter_od %>% 
  group_by(ubigeo_des) %>% 
  summarise(in_flow = sum(cases), .groups = "drop") %>% 
  arrange(desc(in_flow)) ->
  in_flow
```

```{r}
dim(in_flow)
```

```{r}
head(in_flow, 5)
```

## Out-flow

```{r}
inter_od %>% 
  group_by(ubigeo_ori) %>% 
  summarise(out_flow = sum(cases), .groups = "drop") %>% 
  arrange(desc(out_flow)) ->
  out_flow
```

```{r}
dim(out_flow)
```

```{r}
head(out_flow, 5)
```

## Table with in and out-flow for all districts

```{r}
lima_districts <- tibble(ubigeo = unique(od$ubigeo_des))
```

```{r}
lima_districts %>% 
  left_join(in_flow, by = c("ubigeo" = "ubigeo_des")) %>% 
  left_join(out_flow, by = c("ubigeo" = "ubigeo_ori")) %>% 
  replace_na(list(in_flow = 0, out_flow = 0)) ->
  in_out_flow
```

```{r}
in_out_flow
```

## Add population

```{r}
in_out_flow %>% 
  inner_join(pop, by = "ubigeo") %>% 
  select(-c(department, province, district)) -> 
  in_out_flow_pop
```

```{r}
in_out_flow_pop
```
## Mobilization rates

```{r}
quantile_probs <- seq(0, 1, .2)

# in_out_flow_pop %>% 
#   mutate(
#     balance = in_flow - out_flow,
#     in_rate = in_flow / (population - 0.5 * balance),
#     out_rate = out_flow / (population - 0.5 * balance),
#     net_rate = in_rate - out_rate
#   ) -> flow_rates

in_out_flow_pop %>% 
  mutate(
    in_rate = in_flow / (population + in_flow - out_flow),
    out_rate = out_flow / population,
    # net_rate = in_rate - out_rate
  ) -> flow_rates
```

## Categorize rates

```{r}
flow_rates %>% 
  mutate(
    across(
      in_rate:net_rate,
      ~ cut(.x, quantile(.x, quantile_probs), include.lowest = TRUE),
      .names = "{.col}_cat"
    )
  ) %>% 
  mutate(
    net_rate_cat_z = cut(
      net_rate, c(min(flow_rates$net_rate), -0.01, 0.01, max(flow_rates$net_rate)),
      include.lowest = TRUE
    )
  ) -> flow_rates_cat
```

```{r}
flow_rates_cat
```

## Geographic data

```{r}
data("Peru", package = "innovar")
distr_geo <- select(Peru, ubigeo, geometry)
```

```{r}
distr_geo
```

```{r}
distr_geo %>% 
  inner_join(flow_rates_cat, by = "ubigeo") ->
  flow_geo
```

## In-rate plot

```{r}
flow_geo %>% 
  ggplot(aes(fill = in_rate_cat)) + 
  geom_sf(size = 0.1) +
  scale_fill_brewer(palette = "Oranges") -> in_rate_plot

in_rate_plot
```

```{r}
# svg("analysis/figs/in-rate-lima.svg", width = 6, height = 7, pointsize = 12)
# in_rate_plot
# dev.off()
```

## Out-rate plot

```{r}
flow_geo %>% 
  ggplot(aes(fill = out_rate_cat)) + 
  geom_sf(size = 0.1) +
  scale_fill_brewer(palette = "Oranges") -> out_rate_plot

out_rate_plot
```

```{r}
# svg("analysis/figs/out-rate-lima.svg", width = 6, height = 7, pointsize = 12)
# out_rate_plot
# dev.off()
```

## Net-rate plot

```{r}
net_rate_col <- c(
  "[-2.42,-0.01]" = "#D8B365", "(-0.01,0.01]" = "#F5F5F5", 
  "(0.01,7.63]" = "#5AB4AC"
)

flow_geo %>% 
  ggplot(aes(fill = net_rate_cat_z)) + 
  geom_sf(size = 0.1) +
  scale_fill_manual(values = net_rate_col) -> net_rate_plot

net_rate_plot
```

```{r}
# svg("analysis/figs/net-rate-lima.svg", width = 6, height = 7, pointsize = 12)
# net_rate_plot
# dev.off()
```

## District centroids

```{r}
distr_shp_tr <- st_transform(Peru, 2163)
distr_centroids <- st_centroid(distr_shp_tr)
```
## Create od-lines

```{r}
od_lines <- od2line(inter_od, distr_centroids)
od_lines_ow <- od_oneway(od_lines, attrib = "cases")
```
## Interchange data

```{r}
flow_rates %>% 
  select(ubigeo, population, balance) ->
  pop_balance

od_lines_ow %>% 
  inner_join(pop_balance, by = c("ubigeo_ori" = "ubigeo")) %>% 
  inner_join(
    pop_balance, by = c("ubigeo_des" = "ubigeo"), suffix = c("_ori", "_des")
  ) %>% 
  mutate(
    pop_ori = population_ori - 0.5 * balance_ori,
    pop_des = population_des - 0.5 * balance_des,
    interchange = (cases / (0.5 * pop_ori + 0.5 * pop_des)) * 10000
  ) -> od_interchange
```

## Network plot

```{r}
od_interchange <- arrange(od_interchange, interchange)
od_interchange$colour <- viridis(nrow(od_interchange), direction = -1)
od_interchange_min <- filter(od_interchange, interchange > 0)
  
network_plot <- 
  ggplot() +
  geom_sf(data = flow_geo, fill = "white", size = 0.3) +
  geom_sf(
      od_interchange_min, 
      mapping = aes(color = log(interchange), alpha = log(interchange))
    ) +
    scale_color_viridis_c(direction = -1)

network_plot
```

```{r}
# svg("analysis/figs/network-flow-lima.svg", width = 6, height = 7, pointsize = 12)
# network_plot
# dev.off()
```

# Univariate analysis

## Descriptive tables

```{r echo=FALSE}
# statistics = c(
#   "{min}, {max}", "{median} ({p25}, {p75})", "{mean} ({sd})", "{skewness}", 
#   "{kurtosis}"
# )
# stats_labels = c("Range", "Median (IQR)", "Mean (SD)", "Skewness", "Kurtosis")
# digits = rep(2, 9)
```

### Demographics

```{r echo=FALSE}
# dataset |> 
#   dplyr::select(new_cases:work_rate_intra) |> 
#   tbl_summary(
#     statistic = all_continuous2() ~ statistics,
#     digits = all_continuous2() ~ digits,
#     type = everything() ~ "continuous2"
#   ) |> 
#   add_stat_label(label = all_continuous() ~ stats_labels)
```

### Centrality measures

```{r}
centralities_long = dataset |> 
  dplyr::select(district, contains("centrality")) |> 
  tidyr::pivot_longer(
    cols = -district,
    names_to = c("source", "variable", "centrality", "mode"),
    names_sep = "_"
  ) |> 
  tidyr::pivot_wider(
    id_cols = c(district, source),
    names_from = c(variable, centrality, mode),
    values_from = value
  )
```

```{r echo=FALSE}
# centralities_long |> 
#   tbl_summary(
#     by = source,
#     statistic = all_continuous2() ~ statistics,
#     digits = all_continuous2() ~ digits,
#     type = everything() ~ "continuous2",
#     include = -district
#   ) |> 
#   add_stat_label(label = all_continuous() ~ stats_labels)
```

## Exploratory analysis

### Incidence

```{r}
bar_cases = dataset |> 
  group_by(new_cases) |> 
  summarise(count = n(), .groups = "drop") |> 
  ggplot(aes(new_cases, count)) +
  geom_col(width = 1)

bar_cases
```

```{r}
hist_incidence = dataset |> 
  ggplot(aes(incidence)) +
  geom_histogram(bins = 7)

box_incidence = dataset |> 
  ggplot(aes(incidence)) +
  geom_boxplot()

hist_incidence / box_incidence
```

```{r}
hist_incidence_log = dataset |> 
  ggplot(aes(incidence)) +
  geom_histogram(bins = 7)

box_incidence_log = dataset |> 
  ggplot(aes(incidence)) +
  geom_boxplot()

hist_incidence_log / box_incidence_log
```


### Poverty

#### Monetary poverty

```{r}
hist_monetary_poverty = dataset |> 
  ggplot(aes(monetary_poverty)) +
  geom_histogram(bins = 7)

box_monetary_poverty = dataset |> 
  ggplot(aes(monetary_poverty)) +
  geom_boxplot()

hist_monetary_poverty / box_monetary_poverty
```

#### UBN

```{r}
hist_ubn = dataset |> 
  ggplot(aes(ubn)) +
  geom_histogram(bins = 7)

box_ubn = dataset |> 
  ggplot(aes(ubn)) +
  geom_boxplot()

hist_ubn / box_ubn
```

### Pollution

#### CO

```{r}
hist_co = dataset |> 
  ggplot(aes(co)) +
  geom_histogram(bins = 7)

box_co = dataset |> 
  ggplot(aes(co)) +
  geom_boxplot()

hist_co / box_co
```

#### NO2

```{r}
hist_no2 = dataset |> 
  ggplot(aes(no2)) +
  geom_histogram(bins = 7)

box_no2 = dataset |> 
  ggplot(aes(no2)) +
  geom_boxplot()

hist_no2 / box_no2
```

#### O3

```{r}
hist_o3 = dataset |> 
  ggplot(aes(o3)) +
  geom_histogram(bins = 7)

box_o3 = dataset |> 
  ggplot(aes(o3)) +
  geom_boxplot()

hist_o3 / box_o3
```

#### PM25

```{r}
hist_pm25 = dataset |> 
  ggplot(aes(pm25)) +
  geom_histogram(bins = 7)

box_pm25 = dataset |> 
  ggplot(aes(pm25)) +
  geom_boxplot()

hist_pm25 / box_pm25
```

### Intra-communting flow

```{r}
flow_intra_long = dataset |> 
  dplyr::select(district, contains("flow")) |> 
  tidyr::pivot_longer(
    cols = -district,
    names_to = c("source", "variable", "mode"),
    names_sep = "_"
  ) |> 
  tidyr::pivot_wider(
    id_cols = c(district, source),
    names_from = c(variable, mode),
    values_from = value
  )
```

```{r}
hist_flow_intra = flow_intra_long |>  
  ggplot(aes(flow_intra)) +
  geom_histogram(bins = 7) +
  facet_wrap(~source, ncol = 3, scales = "free")

box_flow_intra = flow_intra_long |>  
  ggplot(aes(flow_intra)) +
  geom_boxplot() +
  facet_wrap(~source, ncol = 3, scales = "free")
```

```{r fig.width=6}
hist_flow_intra / box_flow_intra
```

### Centrality measures

#### Degree

##### Out-degree

```{r}
hist_degree_out = centralities_long |>  
  dplyr::select(source, centrality_degree_out) |> 
  ggplot(aes(centrality_degree_out)) +
  geom_histogram(bins = 7) +
  facet_wrap(~source, ncol = 3, scales = "free")

box_degree_out = centralities_long |>  
  dplyr::select(source, centrality_degree_out) |> 
  ggplot(aes(centrality_degree_out)) +
  geom_boxplot() +
  facet_wrap(~source, ncol = 3, scales = "free")
```

```{r fig.width=6}
hist_degree_out / box_degree_out
```

#### Strength

##### Out-strength

```{r}
hist_strength_out = centralities_long |>  
  dplyr::select(source, centrality_strength_out) |> 
  ggplot(aes(centrality_strength_out)) +
  geom_histogram(bins = 7) +
  facet_wrap(~source, ncol = 3, scales = "free")

box_strength_out = centralities_long |>  
  dplyr::select(source, centrality_strength_out) |> 
  ggplot(aes(centrality_strength_out)) +
  geom_boxplot() +
  facet_wrap(~source, ncol = 3, scales = "free")
```

```{r fig.width=6}
hist_strength_out / box_strength_out
```

#### Closeness

##### Out-closeness

```{r}
hist_closeness_out = centralities_long |>  
  dplyr::select(source, centrality_closeness_out) |> 
  ggplot(aes(centrality_closeness_out)) +
  geom_histogram(bins = 7) +
  facet_wrap(~source, ncol = 3, scales = "free")

box_closeness_out = centralities_long |>  
  dplyr::select(source, centrality_closeness_out) |> 
  ggplot(aes(centrality_closeness_out)) +
  geom_boxplot() +
  facet_wrap(~source, ncol = 3, scales = "free")
```

```{r fig.width=6}
hist_closeness_out / box_closeness_out
```

#### Betweenness

```{r}
hist_betweenness_directed = centralities_long |>  
  dplyr::select(source, centrality_betweenness_directed) |> 
  ggplot(aes(centrality_betweenness_directed)) +
  geom_histogram(bins = 7) +
  facet_wrap(~source, ncol = 3, scales = "free")

box_betweenness_directed = centralities_long |>  
  dplyr::select(source, centrality_betweenness_directed) |> 
  ggplot(aes(centrality_betweenness_directed)) +
  geom_boxplot() +
  facet_wrap(~source, ncol = 3, scales = "free")
```

```{r fig.width=6}
hist_betweenness_directed / box_betweenness_directed
```

#### Eigenvector

```{r}
hist_eigenvector_directed = centralities_long |>  
  dplyr::select(source, centrality_eigenvector_directed) |> 
  ggplot(aes(centrality_eigenvector_directed)) +
  geom_histogram(bins = 7) +
  facet_wrap(~source, ncol = 3, scales = "free")

box_eigenvector_directed = centralities_long |>  
  dplyr::select(source, centrality_eigenvector_directed) |> 
  ggplot(aes(centrality_eigenvector_directed)) +
  geom_boxplot() +
  facet_wrap(~source, ncol = 3, scales = "free")
```

```{r fig.width=6}
hist_eigenvector_directed / box_eigenvector_directed
```

#### Page-rank

```{r}
hist_pagerank_directed = centralities_long |>  
  dplyr::select(source, centrality_pagerank_directed) |> 
  ggplot(aes(centrality_pagerank_directed)) +
  geom_histogram(bins = 7) +
  facet_wrap(~source, ncol = 3, scales = "free")

box_pagerank_directed = centralities_long |>  
  dplyr::select(source, centrality_pagerank_directed) |> 
  ggplot(aes(centrality_pagerank_directed)) +
  geom_boxplot() +
  facet_wrap(~source, ncol = 3, scales = "free")
```

```{r fig.width=6}
hist_pagerank_directed / box_pagerank_directed
```

```{r}
ranking = tibble(
  incidence = dataset |> arrange(desc(incidence)) |> pull(district),
  full_flow_intra = dataset |> arrange(desc(full_flow_intra)) |> pull(district),
  study_flow_intra = dataset |> arrange(desc(study_flow_intra)) |> pull(district),
  work_flow_intra = dataset |> arrange(desc(work_flow_intra)) |> pull(district),
  full_centrality_degree_out = dataset |> 
    arrange(desc(full_centrality_degree_out)) |> 
    pull(district),
  study_centrality_degree_out = dataset |> 
    arrange(desc(study_centrality_degree_out)) |> 
    pull(district),
  work_centrality_degree_out = dataset |> 
    arrange(desc(work_centrality_degree_out)) |> 
    pull(district),
  full_centrality_strength_out = dataset |> 
    arrange(desc(full_centrality_strength_out)) |> 
    pull(district),
  study_centrality_strength_out = dataset |> 
    arrange(desc(study_centrality_strength_out)) |> 
    pull(district),
  work_centrality_strength_out = dataset |> 
    arrange(desc(work_centrality_strength_out)) |> 
    pull(district),
  full_centrality_closeness_out = dataset |> 
    arrange(desc(full_centrality_closeness_out)) |> 
    pull(district),
  study_centrality_closeness_out = dataset |> 
    arrange(desc(study_centrality_closeness_out)) |> 
    pull(district),
  work_centrality_closeness_out = dataset |> 
    arrange(desc(work_centrality_closeness_out)) |> 
    pull(district),
  full_centrality_betweenness_directed = dataset |> 
    arrange(desc(full_centrality_betweenness_directed)) |> 
    pull(district),
  study_centrality_betweenness_directed = dataset |> 
    arrange(desc(study_centrality_betweenness_directed)) |> 
    pull(district),
  work_centrality_betweenness_directed = dataset |> 
    arrange(desc(work_centrality_betweenness_directed)) |> 
    pull(district),
  full_centrality_eigenvector_directed = dataset |> 
    arrange(desc(full_centrality_eigenvector_directed)) |> 
    pull(district),
  study_centrality_eigenvector_directed = dataset |> 
    arrange(desc(study_centrality_eigenvector_directed)) |> 
    pull(district),
  work_centrality_eigenvector_directed = dataset |> 
    arrange(desc(work_centrality_eigenvector_directed)) |> 
    pull(district),
  full_centrality_pagerank_directed = dataset |> 
    arrange(desc(full_centrality_pagerank_directed)) |> 
    pull(district),
  study_centrality_pagerank_directed = dataset |> 
    arrange(desc(study_centrality_pagerank_directed)) |> 
    pull(district),
  work_centrality_pagerank_directed = dataset |> 
    arrange(desc(work_centrality_pagerank_directed)) |> 
    pull(district),
)
```

```{r}
readr::write_csv(ranking, "data/processed/ranking.csv")
```

