---
title: "Exploratory data analysis"
output:
  html_document:
    df_print: paged
---

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Packages

```{r}
library(fs)
library(dplyr)
library(readr)
library(tidyr)
library(innovar) # paquete del lab
library(sf) # simple features: trabajar con data geográfica
library(ggplot2)
theme_set(theme_bw())
library(scales)
library(stplanr) # gráficos de redes en mapas
library(viridis) # colores
```
# Load mobilization data

```{r}
interim_path <- path("data", "interim")
file_name <- "mob-lima-metro-od.csv"
file_path <- path(interim_path, file_name)
od <- read_csv(file_path)
glimpse(od)
```

```{r}
length(unique(od$ubigeo_ori))
length(unique(od$ubigeo_des))
```

# Load population data

```{r}
raw_path <- path("data", "raw")
pop_file_name <- "pop-census-2017.csv"
pop_file_path <- path(raw_path, pop_file_name)
pop <- read_csv(pop_file_path)
glimpse(pop)
```

```{r}
#TODO: Intra-district analysis
```

# Inter-distrital analysis

```{r}
od %>% 
  filter(ubigeo_ori != ubigeo_des) ->
  inter_od
```

## Districts that are not a destiny 

```{r}
length(unique(inter_od$ubigeo_des))
setdiff(od$distr_des, inter_od$distr_des)
```

## Districts that are not an origin 

```{r}
length(unique(inter_od$ubigeo_ori))
setdiff(od$distr_ori, inter_od$distr_ori)
```

## In-flow

```{r}
inter_od %>% 
  group_by(ubigeo_des) %>% 
  summarise(in_flow = sum(cases), .groups = "drop") %>% 
  arrange(desc(in_flow)) ->
  in_flow
```

```{r}
dim(in_flow)
```

```{r}
head(in_flow, 5)
```

## Out-flow

```{r}
inter_od %>% 
  group_by(ubigeo_ori) %>% 
  summarise(out_flow = sum(cases), .groups = "drop") %>% 
  arrange(desc(out_flow)) ->
  out_flow
```

```{r}
dim(out_flow)
```

```{r}
head(out_flow, 5)
```

## Table with in and out-flow for all districts

```{r}
lima_districts <- tibble(ubigeo = unique(od$ubigeo_des))
```

```{r}
lima_districts %>% 
  left_join(in_flow, by = c("ubigeo" = "ubigeo_des")) %>% 
  left_join(out_flow, by = c("ubigeo" = "ubigeo_ori")) %>% 
  replace_na(list(in_flow = 0, out_flow = 0)) ->
  in_out_flow
```

```{r}
in_out_flow
```

## Add population

```{r}
in_out_flow %>% 
  inner_join(pop, by = "ubigeo") %>% 
  select(-c(department, province, district)) -> 
  in_out_flow_pop
```

```{r}
in_out_flow_pop
```
## Mobilization rates

```{r}
quantile_probs <- seq(0, 1, .2)

# in_out_flow_pop %>% 
#   mutate(
#     balance = in_flow - out_flow,
#     in_rate = in_flow / (population - 0.5 * balance),
#     out_rate = out_flow / (population - 0.5 * balance),
#     net_rate = in_rate - out_rate
#   ) -> flow_rates

in_out_flow_pop %>% 
  mutate(
    in_rate = in_flow / (population + in_flow - out_flow),
    out_rate = out_flow / population,
    # net_rate = in_rate - out_rate
  ) -> flow_rates
```

## Categorize rates

```{r}
flow_rates %>% 
  mutate(
    across(
      in_rate:net_rate,
      ~ cut(.x, quantile(.x, quantile_probs), include.lowest = TRUE),
      .names = "{.col}_cat"
    )
  ) %>% 
  mutate(
    net_rate_cat_z = cut(
      net_rate, c(min(flow_rates$net_rate), -0.01, 0.01, max(flow_rates$net_rate)),
      include.lowest = TRUE
    )
  ) -> flow_rates_cat
```

```{r}
flow_rates_cat
```

## Geographic data

```{r}
data("Peru", package = "innovar")
distr_geo <- select(Peru, ubigeo, geometry)
```

```{r}
distr_geo
```

```{r}
distr_geo %>% 
  inner_join(flow_rates_cat, by = "ubigeo") ->
  flow_geo
```

## In-rate plot

```{r}
flow_geo %>% 
  ggplot(aes(fill = in_rate_cat)) + 
  geom_sf(size = 0.1) +
  scale_fill_brewer(palette = "Oranges") -> in_rate_plot

in_rate_plot
```

```{r}
# svg("analysis/figs/in-rate-lima.svg", width = 6, height = 7, pointsize = 12)
# in_rate_plot
# dev.off()
```

## Out-rate plot

```{r}
flow_geo %>% 
  ggplot(aes(fill = out_rate_cat)) + 
  geom_sf(size = 0.1) +
  scale_fill_brewer(palette = "Oranges") -> out_rate_plot

out_rate_plot
```

```{r}
# svg("analysis/figs/out-rate-lima.svg", width = 6, height = 7, pointsize = 12)
# out_rate_plot
# dev.off()
```

## Net-rate plot

```{r}
net_rate_col <- c(
  "[-2.42,-0.01]" = "#D8B365", "(-0.01,0.01]" = "#F5F5F5", 
  "(0.01,7.63]" = "#5AB4AC"
)

flow_geo %>% 
  ggplot(aes(fill = net_rate_cat_z)) + 
  geom_sf(size = 0.1) +
  scale_fill_manual(values = net_rate_col) -> net_rate_plot

net_rate_plot
```

```{r}
# svg("analysis/figs/net-rate-lima.svg", width = 6, height = 7, pointsize = 12)
# net_rate_plot
# dev.off()
```

## District centroids

```{r}
distr_shp_tr <- st_transform(Peru, 2163)
distr_centroids <- st_centroid(distr_shp_tr)
```
## Create od-lines

```{r}
od_lines <- od2line(inter_od, distr_centroids)
od_lines_ow <- od_oneway(od_lines, attrib = "cases")
```
## Interchange data

```{r}
flow_rates %>% 
  select(ubigeo, population, balance) ->
  pop_balance

od_lines_ow %>% 
  inner_join(pop_balance, by = c("ubigeo_ori" = "ubigeo")) %>% 
  inner_join(
    pop_balance, by = c("ubigeo_des" = "ubigeo"), suffix = c("_ori", "_des")
  ) %>% 
  mutate(
    pop_ori = population_ori - 0.5 * balance_ori,
    pop_des = population_des - 0.5 * balance_des,
    interchange = (cases / (0.5 * pop_ori + 0.5 * pop_des)) * 10000
  ) -> od_interchange
```

## Network plot

```{r}
od_interchange <- arrange(od_interchange, interchange)
od_interchange$colour <- viridis(nrow(od_interchange), direction = -1)
od_interchange_min <- filter(od_interchange, interchange > 0)
  
network_plot <- 
  ggplot() +
  geom_sf(data = flow_geo, fill = "white", size = 0.3) +
  geom_sf(
      od_interchange_min, 
      mapping = aes(color = log(interchange), alpha = log(interchange))
    ) +
    scale_color_viridis_c(direction = -1)

network_plot
```

```{r}
# svg("analysis/figs/network-flow-lima.svg", width = 6, height = 7, pointsize = 12)
# network_plot
# dev.off()
```



